---
title: "APW_data_explore1"
author: "Sylvia Soltyk"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = FALSE}
library(lubridate)
library(tidyr)
library(dplyr)
library(naniar)
library(visdat)
library(simputation)
library(ggplot2)
library(scales)
library(viridis)
library(tseries)
library(dse)
```

#Load data:
(There were duplicaed dates which I removed manually in the csv)
```{r}
APW_data <- read.csv("/Users/sylviasoltyk/Dropbox/PhD/Data/Mecardo_wheat/APW_analysis/APW.csv")
APW_data$Date <- dmy(APW_data$Date)
```

#Cleaning data: examining for outliers via histograms and box plots
```{r}
hist(APW_data$Adelaide)
hist(APW_data$Geelong)
hist(APW_data$Kwinana)
hist(APW_data$Port.Kembla)
hist(APW_data$Port.Lincoln)

boxplot(APW_data[,-1]) 
```

#Cleaning data: identifying where missing data is
this shows that there are NA's in the data (missing values) # Adelaide 17, Geelong 10, Kwinana 103, Port Kembla 38, Port Lincoln 38 
```{r}
summary(APW_data) 
NAind_Adelaide <- which(is.na(APW_data$Adelaide))
NAind_Geelong <- which(is.na(APW_data$Geelong))
NAind_Kwinana <- which(is.na(APW_data$Kwinana))
NAind_Port.Kembla <- which(is.na(APW_data$Port.Kembla))
NAind_Port.Lincoln <- which(is.na(APW_data$Port.Lincoln))
```

proportion of missing values in series. Adeliaide 0.006706114, Geelong 0.003944773, Kwinana 0.04063116, Port Kembla 0.01499014, Port Lincoln 0.01499014
```{r}
prop_miss(APW_data$Adelaide)
prop_miss(APW_data$Geelong)
prop_miss(APW_data$Kwinana)
prop_miss(APW_data$Port.Kembla)
prop_miss(APW_data$Port.Lincoln)
```

find longest stretch of NA's in data # Adelaide 2, Geelong 1, Kwinana 4, Port Kembla 18, Port Lincoln 6 
```{r}
longestNAstrech <- function(x) {
  with(rle(is.na(x)), max(lengths[values]))  
}
longestNAstrech(APW_data$Adelaide) 
longestNAstrech(APW_data$Geelong) 
longestNAstrech(APW_data$Kwinana) 
longestNAstrech(APW_data$Port.Kembla) 
longestNAstrech(APW_data$Port.Lincoln) 

```

visualisations of missingness
```{r}
vis_miss(APW_data[,-1])
```

#Imputing missing values
method: last observation carried forward (LOCF)
```{r}
APW_LOCF <- APW_data %>% fill(Adelaide, Geelong, Kwinana, Port.Kembla, Port.Lincoln)
```

exploring the completed data set (no missing values) via a boxplot
```{r}
boxplot(APW_LOCF[,-1]) 
```

creating a plot of the completed data

```{r}
ggplot(data=APW_LOCF, aes(x=Date)) + 
  geom_line(aes(y = Adelaide, color = "Adelaide")) + 
  geom_line(aes(y = Geelong, color="Geelong")) +
  geom_line(aes(y = Kwinana, color="Kwinana"))+
  geom_line(aes(y = Port.Kembla, color="Port.Kembla")) +
  geom_line(aes(y = Port.Lincoln, color="Port.Lincoln")) +
  scale_y_continuous(limits = c(150,500), breaks=seq(150,500,by=50)) + 
  scale_x_date(date_breaks = '6 months' , date_labels = "%b-%y") +
  theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  labs(x = "Date",
       y = "AU$/t") 
```

#checking for stationarity (doesn't look stationary according to the plots)
augmented dickey fuller test. Alternate hypothesis is stationarity. Result for all: dnr - non stationary
```{r}
adf.test(APW_LOCF$Adelaide) 
adf.test(APW_LOCF$Geelong) 
adf.test(APW_LOCF$Kwinana) 
adf.test(APW_LOCF$Port.Kembla) 
adf.test(APW_LOCF$Port.Lincoln) 
```

acf - seems to confirm non staitonarity
```{r}
acf(APW_LOCF$Adelaide)
acf(APW_LOCF$Geelong)
acf(APW_LOCF$Kwinana)
acf(APW_LOCF$Port.Kembla)
acf(APW_LOCF$Port.Lincoln)
```

#Log difference (%) to create a stationary data set
taking first log differences
```{r}
APW_LOCF_diff<-data.frame(APW_LOCF$Date[-1], 100*diffLog(APW_LOCF$Adelaide), 100*diffLog(APW_LOCF$Geelong), 100*diffLog(APW_LOCF$Kwinana), 100*diffLog(APW_LOCF$Port.Kembla), 100*diffLog(APW_LOCF$Port.Lincoln))
names(APW_LOCF_diff)<-names(APW_LOCF)
```

just having a look at the data in a boxplot
```{r}
boxplot(APW_LOCF_diff[,-1]) 
```

plot the log differences data
```{r}
ggplot(data=APW_LOCF_diff, aes(x=Date)) + 
    geom_line(aes(y = Adelaide, color = "Adelaide")) + 
    geom_line(aes(y = Geelong, color="Geelong")) +
    geom_line(aes(y = Kwinana, color="Kwinana"))+
    geom_line(aes(y = Port.Kembla, color="Port.Kembla")) +
    geom_line(aes(y = Port.Lincoln, color="Port.Lincoln")) +
    scale_y_continuous(limits = c(-25,25), breaks=seq(-25,25,by=5)) + 
    scale_x_date(date_breaks = '6 months' , date_labels = "%b-%y") +
    theme(axis.text.x = element_text(angle = 45, hjust=1)) +
    labs(x = "Date",
         y = "% daily log return AU$/t") 
```

check stationarity using Dickey-Fuller test. Result for all: reject, is stationary
```{r}
adf.test(APW_LOCF_diff$Adelaide) 
adf.test(APW_LOCF_diff$Geelong) 
adf.test(APW_LOCF_diff$Kwinana) 
adf.test(APW_LOCF_diff$Port.Kembla) 
adf.test(APW_LOCF_diff$Port.Lincoln) 
```

